%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Implementation.tex                                  %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Andre C. Marta                                           %
%     Last modified :  2 Jul 2015                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Background concepts}
\label{chapter:bc}

Before starting explaining what was done, we introduce the main concepts and tools required.

We will start by introducing the field of Optimal Control Problems: defining its key elements, deducing the dynamic programming principle, respective dynamic programming equation and the verification technique. Secondly, we will explain how they are related with the field of optimal stopping problems as well as how we can find an optimal stopping time. Lastly, we will deduce the general solution of the (standard) optimal problems that we will face along this work, taking into account its specific financial context.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Stochastic Control Problems}
\label{section:control}

\subsubsection{Key elements}

We need to set the key elements related to stochastic control problems. Among these, it is included:

\textbf{Time Horizon:}
in the context of investment-consumption problems (as this work is), an infinite time-horizon is considered: $[0,\infty)$. However one may also consider a finite horizon: $[0,T] \ T \in(0,\infty)$, as in the case where a situation is valid until an expiration date $T$; or and indefinite horizon: $[0,\tau]$ for some $\tau$ stopping time, as in the case where a situation is valid until a certain exit time $\tau$.


\textbf{Controlled State Process:}  
a stochastic process which describes the state of the event that we are interested on and that it is often given as the solution of a stochastic differential equation (SDE) of the form
\begin{equation}
	d X_t=b(t,X_t,U_t)dt + \sigma (t,X_t,U_t)dW_t, \ X_0=x
	\label{bc_cp}
\end{equation}
where $\{U_t: \ t \in T_u\}$ is a control as defined hereunder; $\{W_t: \ t\geq 0\}$ corresponds to a standard Brownian Motion and $b$ and $\sigma$ are functions that satisfy the global Lipschitz and linear growth conditions (also known by Itô conditions), respectively given by
\begin{align}
\exists K \in (0,\infty) \  \forall t \in [0,\infty) \ \forall u \in \mathds{U} \ \forall x,y \in\mathds{R}^n: \hspace{3mm} \nonumber &\\
|b(t,x,\alpha)-b(t,y,\alpha)| + || \sigma (t,x,\alpha)- \sigma(t,y,\alpha)|| &\leq K |x-y| \label{bc_cond1} \\
|b(t,x,\alpha)|^2+|| \sigma (t,x,\alpha)|| &\leq K^2 (1+|x|^2).  \label{bc_cond2}
\end{align}


\textbf{Control process:}
a stochastic process chosen to influence the state of the system. Considering the SDE \eqref{bc_cp}, $U=\{U_t: \ t \in T_u\}$ is a control that influences the drift and the diffusion of the SDE and that takes values on \textit{control space} $\mathds{U} \in \mathds{R}^m$, with $T_u$ representing the \textit{support space}.


\textbf{Admissible Controls:}
control processes that verify certain conditions impose by the context of the problem. More it will be explained on Section \ref{}.


\textbf{Cost/reward function:} 
denoted as $J(x,U)$, it identifies the costs/rewards of a system with an initial state $x$ and a control process $U$.


\textbf{Value Function:}
denoted as $V(x)$, it corresponds to the minimum/maximum possible cost/reward of the system over all admissible controls. Hence, when dealing with a cost function $V$ is such that $V(x)= \underset{U\in \mathds{U}}{\inf} J(x,U)$. On the other side, when dealing with a reward function (which will be our case) $V$ is such that $V(x)= \underset{U\in \mathds{U}}{\sup} J(x,U)$.


Therefore we can now state the primary goal of a stochastic control problem: for a given initial state $x$, we want to find the control process that lead to the optimized value function $V(x)$.

How one is able to derive this optimal control process? That's what will be explained in the next section, in which we show the main tools used on this sort of problems and proof their correctness.

\subsubsection{Dynamic programming principle}

We will follow the explanation made in \cite{ross}. Hence we assume an infinite horizon time $[0,\infty)$ and a controlled state process as in \eqref{bc_cp} defined on the probability space $(\Omega,\mathcal{F}, \mathds{P})$ associated to the underlying Brownian Motion, on which $\Omega$ corresponds to its state space, $\mathcal{F}=\{\mathcal{F}_t, \ t\geq0 \}$ the natural filtration associated to it and the $\mathds{P}$ the probability measure. Considering a unidimensional controlled state space and under these assumptions, \eqref{bc_cp} takes the form of
\begin{equation}
 	d X_t=b(X_t,U_t)dt + \sigma (X_t,U_t)dW_t, \ X_0=x\in \mathds{R},
 \label{bc_cp2}
\end{equation} 
with $U$ being the control process defined in the control space $\mathds{U}=mathds{R}$.

Due to the close connection to our problem, we will restrict our attention to Markov controls.
\begin{defi}
	A \textit{Markov control process} is a stochastic process that takes the
	form $U_t=u(X_t) \ \forall t \geq 0$ for some $u: \mathds{R} \rightarrow \mathds{U}=\mathds{R}$.
\end{defi}

Also, assume that the state process $X \equiv X_x^u$ \eqref{bc_cp2} associated with the Markov control function $u$ is a time-homogeneous Markov process with generator $\mathcal{A} \equiv \mathcal{A}^u$, that is
$$\mathcal{A}=\lim_{t\downarrow 0} \frac{\mathds{E}^{X_0=x}(f(X_t))-f(x)}{t}=b(x,u(x)) f'(x)+\frac{1}{2}\sigma^2(x,u(x))f''(x),$$
with $f \in C^2(\mathds{R})$. Note that $X$ is uniquely determined by the initial state $x$, the Markov control function $u$ and functions $b$ and $\sigma$, which verify conditions \eqref{bc_cond1} and \eqnref{bc_cond2}. If functions $b$ and $\sigma$ are fixed (assumption that will hold from now on), $X$ will be completely determined by $x$ and $u$.

We now define the cost function $J$ as
\begin{equation}
	 J(x,u)=\mathds{E}\left[ \int^\infty_0 e^{-\gamma s} g(X_s,u(X_s)) \ ds | X_0=x \right],
	 \label{bc_J}
\end{equation}
where $\mathds{E} \equiv \mathds{E}^u$, and the respective value function as $V$ as
\begin{equation}
	V(x)=\inf_{u \in \mathds{U}} J(x,u),
	\label{bc_V}
\end{equation}
where the infimum is taken over all the Markov control functions.

For the sake of simplicity, we will ignore the admissibility restrictions and consider that always exists a Markov control function $u^*$ such that $V(x)=J(x,u^*) \ \forall x$ and the conditional expectation as in \eqref{bc_J} to be $\mathds{E}\left[\ . \ | X_0=x \right] \equiv \mathds{E}\left[\ . \ \right]$. Also, denote $X^*$ to be the controlled state process associated with $u^*$. Thus it holds
\begin{equation}
V(x)=J(x,u^*)=\mathds{E}\left[ \int^t_0 e^{-\gamma s} g(X^*_s,u^*(X^*_s)) \ ds\right] + \mathds{E}\left[ \int^\infty_t e^{-\gamma s} g(X^*_s,u^*(X^*_s)) \ ds\right]
	\label{bc_V2}
\end{equation}

Manipulating the expression $\mathds{E}\left[ \int^\infty_t e^{-\gamma s} g(X^*_s,u^*(X^*_s)) \ ds\right]$ we obtain
\begin{align}
\mathds{E}\left[ \int^\infty_t e^{-\gamma s} g(X^*_s,u^*(X^*_s)) \ ds\right]
=& e^{-\gamma t}\mathds{E}\left[ \int^\infty_t e^{-\gamma (s-t)}g(X^*_s,u^*(X^*_s)) \ ds\right] \nonumber \\
&\underset{v:=s-t}{=}e^{-\gamma t}\mathds{E}\left[ \int^\infty_0 e^{-\gamma v}g(X^*_{v+t},u^*(X^*_{v+t})) \ ds\right] \nonumber \\
= & e^{-\gamma t}\mathds{E}\left[ \int^\infty_0 e^{-\gamma v}g(\bar{X}^*_{v+t},u^*(\bar{X}^*_{v+t})) \ ds\right]. \label{bc_ded}
\end{align}

In the last step we introduced the process $\{\bar{X}^*_v=X^*_{v+t}: \ v \geq 0\}$ for some $t\geq 0$. Since $X^*$ is a time-homogeneous Markov process, $\bar{X}^*$ also it is. Therefore its also determined by function $b$ and $\sigma$, the same control function $u^*$ and its the initial state. Furthermore, if we consider $ \bar{X}^*_0=X^*_0=x$ (note that here $t=0$) we have that the conditional expected value on \eqref{bc_ded} is
\begin{align}
	\mathds{E}\left[ \int^\infty_0 e^{-\gamma v}g(\bar{X}^*_{v+t},u^*(\bar{X}^*_{v+t})) \ ds\right]
	=& \mathds{E}\left[ \int^\infty_0 e^{-\gamma v}g(\bar{X}^*_{v+t},u^*(\bar{X}^*_{v+t})) \ ds | \bar{X}^*_0=X^*_0=x \right] \nonumber \\
	&\underset{s=v+t}{=}\mathds{E}\left[ \int^\infty_0 e^{-\gamma v}g(X^*_s,u^*(X^*_s)) \ ds  \right] \nonumber \\
	=&J(x,u^*) \nonumber \\
	=&V(x). \label{bc_ded2}
\end{align}

Following a similar approach considering the random starting state $\bar{X}^*_0=X^*_t$, it follows
\begin{align}
\mathds{E}\left[ \int^\infty_0 e^{-\gamma s}g(X^*_s,u^*(X^*_s)) \ ds  \right]
=& e^{-\gamma t} \mathds{E}\left[ \int^\infty_0 g(\bar{X}^*_{v},u^*(\bar{X}^*_{v})) \ dv | \bar{X}^*_0=X^*_0=x \right] \nonumber \\
=& e^{-\gamma t}  \mathds{E}\left[ \mathds{E}\left[ \int^\infty_0 g(\bar{X}^*_{v},u^*(\bar{X}^*_{v})) \ dv | \bar{X}^*_0=x | \bar{X}^*_0  \right] \right] \nonumber \\
% =& e^{-\gamma t}  \mathds{E}\left[ J(\bar{X}_0^*,u^*)  \right] \\
% =&  e^{-\gamma t}  \mathds{E}\left[ J(X_t^*,u^*)  \right] \\
=& e^{-\gamma t}  \mathds{E}\left[ V(X_t^*)  \right]. \label{bc_ded3}
\end{align}

On the first step we used the fact $\{ \bar{X}_t^*, \ t\geq 0\}\overset{D}{=} \{ X_t^*, \ t\geq 0\}$. On the second step we used the Tower Rule, conditioning to $\bar{X}^*_0=X^*_0=x$. On the last step the result on \eqref{bc_ded2} was used.

Replacing \eqref{bc_ded3} on \eqref{bc_V2}, we obtain
\begin{equation}
 V(x)= \mathds{E}\left[ \int^t_0 e^{-\gamma s} g(X^*_s,u^*(X^*_s)) \ ds\right] + e^{-\gamma t}  \mathds{E}\left[ V(X_t^*)  \right].
 \label{bc_V3}
\end{equation}

Now consider $u$ an arbitrary Markov control function, as well as the associated controlled process $X^u$, and $\hat{u}$ a Markov control function such that
$$\hat{u}=\begin{cases}
u, \ 0\leq s <t \\
u^*, \ s \geq t
\end{cases}.$$

Following a similar strategy used to derive \eqref{bc_V3} we obtain
\begin{equation}
V(x) \leq \mathds{E}\left[ \int^t_0 e^{-\gamma s} g(X^u_s,u^*(X^u_s)) \ ds\right] + e^{-\gamma t}  \mathds{E}\left[ V(X_t^u)  \right].
\label{bc_V4}
\end{equation}

Combining results \eqref{bc_V3} and \eqref{bc_V4} we obtain the \textit{Dynamic Programming Principle} corresponding to 
\begin{equation}
V(x)=\inf_{u \in \mathds{U}} \mathds{E}\left[ \int^t_0 e^{-\gamma s} g(X^u_s,u^*(X^u_s)) \ ds+ e^{-\gamma t} V(X_t^u)  \right],
\label{bc_V5}
\end{equation}
where $\mathds{U}$ is considered to be the set of all (admissible) Markov control functions. It states that the optimal strategy found is either being calculated on each of the time intervals $[0,t)$ and $[t,\infty)$ or on the whole time interval $[0,\infty)$, leading to the same result. The same statement holds for any partition of the time interval.



\subsubsection{Dynamic Programming Equation}

We will explain how one can find the optimal control function $u^*$ using the Dynamic Programming Principle. A similar approach as in \cite{ross} will be followed (so as the previous notation).

Consider an arbitrary Markov control $u$ with corresponding state process $X$ with initial state $X_0=x$ and generator $\mathcal{A}$. Applying Itô's Lemma to the function $f(t,x)=e^{\gamma t}V(x)$ (assumed to be $C(\mathds{R}^2)$ and integrating we obtain
\begin{equation}
e^{\gamma t}V(X_t)-V(X_0)=\int^t_0  e^{\gamma t} \left( - \gamma V(X_s))+\mathcal{A}V(X_s)\right) \ ds + \sigma \int^t_0 \frac{\partial}{\partial x}(e^{-\gamma s}V(x))\bigg\rvert_{x=X_s} dW_s.
\label{bc_dpe1}
\end{equation}

Noting that $X_0=x$ and taking the expect value on both sides we obtain
\begin{equation}
\mathds{E} \left[  e^{\gamma t}V(X_t) \right]=V(x) + \mathds{E} \left[ \int^t_0  e^{\gamma t} \left( - \gamma V(X_s))+\mathcal{A}V(X_s)\right) \ ds \right],
\label{bc_ded4}
\end{equation}
where we used the fact that $\int^t_0 \frac{\partial}{\partial x}(e^{-\gamma s}V(x))\bigg\rvert_{x=X_s} dW_s$ is a martingale - since $\{ \frac{\partial}{\partial x}(e^{-\gamma s}V(x)) \in B \} \in \mathcal{F}_s \ \forall B$ Borel-set and 
$\int^t_0 \left( \frac{\partial}{\partial x}(e^{-\gamma s}V(x)) \right)^2  ds
=\frac{1}{2\gamma}\left( 1 - e^{-2\gamma s} \left(\frac{\partial}{\partial x}V(x)\right)^2 \right)<\infty \forall V(x),t<\infty$ - and, hence, it's expected value is 0.

Using results deduced in \eqref{bc_V4} and \ref{bc_ded4} and by the dynamic programming principle, we obtain
\begin{align}
V(x)+\mathds{E} \left[ \int^t_0  e^{\gamma s} \left( - \gamma V(X_s))+\mathcal{A}V(X_s)\right) \ ds \right] &\geq V(x) - \mathds{E} \left[ \int^t_0  e^{\gamma s} g(X_s^u,u(X_s^u)) ds \right] \nonumber \\ 
\Rightarrow \ \mathds{E} \left[ \int^t_0  e^{\gamma s} \left( - \gamma V(X_s))+\mathcal{A}V(X_s)+g(X_s^u,u(X_s^u))\right) \ ds \right] &\geq 0.
\label{bc_ded5}
\end{align}

Assuming that exists an optimal control function $u^*$ it is such that
$\mathds{E} \left[ \int^t_0  e^{\gamma s} \left( - \gamma V(X^*_s))+\mathcal{A}^*V(X^*_s)+g(X^*_s,u^*(X_s^*))\right) \ ds \right]=0$ holds, where $X^*$ and $\mathcal{A}^*$ are the corresponding controlled process and generator, respectively.
 
Now, we divide the expression in \eqref{bc_ded5} by $t$ and we take its limit as $t \downarrow 0$. Taking into account that $X_t \rightarrow X_0=x \ \Rightarrow u(X_t) \rightarrow u(x)$ and considering that the different functions are smooth such that we are able to 
interchange the integral with the expectation (using Fubini's theorem) it follows
\begin{equation}
 - \gamma V(x)+\mathcal{A}V(x)+g(x,u(x))\geq0.
 \label{bc_ded6}
\end{equation}

Considering now the optimal control function $u^*$, $\eqref{bc_ded6}$ corresponds to
\begin{equation}
- \gamma V(x)+\mathcal{A^*}V(x)+g(x,u^*(x))=0.
\label{bc_ded7}
\end{equation}

Summarizing \eqref{bc_ded6} and \eqref{bc_ded7}, we obtain the \textit{dynamic programming equation} (DPE) that is given by
\begin{equation}
	 \inf_{u \in \mathds{U}} \{  - \gamma V(x)+\mathcal{A}V(x)+g(x,u(x)) \}=0,
	 \label{bc_eq}
\end{equation}
where the infimum is takenover all Markov control functions.
Observe that the DPE establishes a map between $x \in \mathds{S}$ (an initial observation in the set of states) and $u \in \mathds{U}$ (corresponding to the optimal control function). Therefore we have that by finding a solution to the DPE, we find the optimal control.

However many assumptions that we made along the explanation - such as the smoothness of the value function $V$ or necessary/sufficient conditions that must hold in order to assure the existence of an optimal control - are hardly observed. Nevertheless there are two main approaches that can be used to obtain a solution to the DPE, while verifying all necessary assumptions: using the theory of \textit{viscosity solutions}, that we will not go further here (for further details check \cite{ross} and \cite{oksendal:book}) or using the \textit{verification} technique, whose main idea we will explain.



\subsubsection{Verification}
 
The verification is seen as a backwards technique. Instead of solving the optimal control problem using the correspondent DPE, we suppose a solution was found and we show that this solution satisfy the admissibility conditions and that it cost/reward corresponds to the value function associated to the problem in hands. Its a very useful method since if we can find a solution to the DPE, then that solution gives us what we want, an optimal control.


Once again (but now in order to emphasize that we want to find an ODE's solution) our optimal control problem might be stated as: we want to find  $\phi \in C^2(\mathds{R})$ that satisfies the DPE, that is,
\begin{equation}
  \inf_{\alpha \in \mathds{U}} \{  - \gamma \phi(x)+\mathcal{A^\alpha}\phi(x)+g(x,\alpha) \}=0 \ x\in \mathds{R}.
  \label{bc_eq2}
\end{equation}

Using the verification technique we are able to find a relationship between the value function $V$ and our solution $\phi$, that a priori doesn't exist.
Therefore we assume that exists a function $\phi \in C^2(\mathds{R})$ that satisfies \eqref{bc_eq2}, we fix $x \in \mathds{R}$ and the correspondent unique optimal control function $\alpha_x^* \in \mathds{U}$ that always exists and satisfies
\begin{equation}
 \alpha_x^*= \arg \min_{\alpha \in \mathds{U} }  \{  - \gamma \phi(x)+\mathcal{A^\alpha}\phi(x)+g(x,\alpha) \}  \ x\in \mathds{R},
 \label{bc_a}
\end{equation}
which corresponds to a map between $x \in \mathds{R}$ and $\alpha_x^* \in \mathds{U}$. From now on this map will be denoted by $u^*: \mathds{R} \rightarrow \mathds{U}$ and we will show that it defines an optimal Markov control function.

However, since any Markov control process is an admissible control process, we need to define the class of admissible control processes. As stated in \cite{ross} we have:
\begin{defi}
	A stochastic process $U= \{ U_t, \ t \geq 0 \}$ is an admissible control process if:
	\begin{enumerate}
		\item $U$ is $\{ \mathcal{F}_t\}$-adapted;
		\item $U_t \in \mathds{U} \ \forall t \geq 0$;
		\item The SDE in \eqref{bc_cp} has a unique solution;
		\item The process $\int^t_0 e^{-\gamma s} \phi^\prime(X_s)\sigma(X_s,U_s) dW_s$ is a martingale;
		\item $e^{-\gamma t} \mathds{E}[V(X_t)] \rightarrow 0$ as $t \rightarrow \infty$. 
	\end{enumerate}
\end{defi}

In order to show that $\phi$ corresponds to the value function $V$, consider $\mathds{A}$ to be the set of admissible controls, that are not necessarily Markov. The cost and value functions are defined in a similar way as before, by taking the infimum over all admissible controls:
\begin{align*}
	J(x,U)&=\mathds{E} \left[ \int^\infty_0 e^{-\gamma t}g(X_t,U_t)dt\right]  \\
	V(x)&=\inf_{U \in \mathds{A}} J(x,U)
\end{align*}

The verification technique uses the same reasoning as used on the DPE. We define an associated SDE by $e^{-\gamma t}\phi(X_t^*)$ and using Itô's lemma we integrate from 0 to $t$, obtaining
\begin{equation}
e^{-\gamma t}\phi(X_t^*)=\phi(x)+\int^t_0 e^{-\gamma s}[-\gamma \phi(X_s^*)+\mathcal{A^*}\phi(X^*_s)] ds+\int^t_0 e^{-\gamma s} \phi^\prime(X^*_s)\sigma(X^*_s,U^*_s) dW_s,
\label{bc_bah}
\end{equation}
where $\{X_t^*, \ t \geq 0\}$ denotes the process $X$ associated with the optimal control function $\alpha^*_x$. This calculation is possible since $\phi \in C^2(\mathds{R}$.

Now, adding $\int^\infty_0 e^{-\gamma s}g(X_s^*,U_s^*)ds$ and then taking the expected value on both sides of \eqref{bc_bah}, it follows that
\begin{equation}
\mathds{E} \left[ \int^\infty_0 e^{-\gamma s}g(X_s^*,U_s^*)ds \right] + \mathds{E} \left[ e^{-\gamma t}\phi(X_t^*) \right]=
\phi(x)+ \mathds{E} \left[ \int^t_0 e^{-\gamma s}(-\gamma \phi(X_s^*)+\mathcal{A^*}\phi(X^*_s)+g(X_s^*,U_s^*)) ds \right],
\label{bc_bah2}
\end{equation}
where we used the fact that the integral $\int^t_0 e^{-\gamma s} \phi^\prime(X^*_s)\sigma(X^*_s,U^*_s) dW_s$ is a martingale and thus its expected value is 0.


Recalling that $\phi$ is the solution of \eqref{bc_a}, then by the DPE $\gamma \phi(x)+\mathcal{A}^\alpha \phi(x)+g(x,\alpha)=0$. (RECTIFICAR!)

Since we are considering the set of all admissible control processes, when taking $t \rightarrow \infty$, it follows that $e^{-\gamma t} \mathds{E}[V(X_t)] \rightarrow 0$.

Hence, in the limit, \eqref{bc_bah2} simplifies to
\begin{equation}
	\phi(x)=\mathds{E} \left[ \int^\infty_0 e^{-\gamma s}g(X_s^*,U_s^*)ds \right]=J(x,U^*)
	\label{bc_bah3}
\end{equation}

On the other side, when considering an arbitrary admissible control process $U \in \mathds{A}$, it follows by the DPE that $\phi(X_s)+\mathcal{A}^\alpha \phi(X_s)+g(X_s,U_S \geq 0 \ \forall s \in [0, \infty)$, which taken into the limit as done before ( $t \rightarrow \infty$) follows that
\begin{equation}
\phi(x) \leq \mathds{E} \left[ \int^\infty_0 e^{-\gamma s}g(X_s,U_s)ds \right]=J(x,U).
\label{bc_bah4}
\end{equation}

Joining both results \eqref{bc_bah3} and \eqref{bc_bah4}, we obtain
\begin{equation}
\phi(x) \leq \mathds{E} \left[ \int^\infty_0 e^{-\gamma s}g(X_s,U_s)ds \right]=J(x,U) \ \underset{\forall U \in \mathds{A}}{\Rightarrow} \ \phi(x) \leq \inf_{U \in \mathds{A}} J(x,U)=V(x).
\label{bc_bah5}
\end{equation}

However, since in \eqref{bc_bah3} we proved that $\phi(x)=J(x,U^*)$ and by the rightmost relation in \eqref{bc_bah5}, which states that $\phi(x)=J(x,U^*)=V(x)$, it follows that $U^*$ is an optimal control.


\section{Optimal Stopping Problems}

Having treated the general context of optimal control problems, now we specify in the field of optimal stopping problems. While on an optimal control problem our goal was to find an optimal control, on an optimal stopping problem our goal is to find the optimal time to achieve a certain purpose, optimizing our cost/reward function. Hence, considering a optimal control problem where the set of admissible controls is taken to have controls that for each state $x$ find the optimal (time to make an) action (if we should continue or stop), we have that this problem coincides with an optimal stopping problem.


\subsubsection{Settings}
 The general formulation is quite identical to the one stated for optimal control problems (now considering an optimal stopping time as the control function), as it will be seen hereunder. 
 For the sake of simplicity and the close relation with what will be developed in this thesis, we will focus on the unidimensional problem.
 
 
 We start by considering the probability space $(\Omega,\mathcal{F}, \mathds{P})$ associated to the underlying Brownian Motion $W$, on which $\mathcal{F}=\{\mathcal{F}_t, \ t\geq0 \}$ corresponds to its natural filtration and a stochastic process $X=\{ X_t, \ t \geq0 \}$ with state space $\mathds{S}=\mathds{R}$, which evolves according to the following SDE
 \begin{equation}
 d X_t=b(X_t)dt + \sigma (X_t)dW_t, \ X_0=x\in \mathds{R},
 \label{bc_cp2}
 \end{equation} 
 where $b$ and $\sigma$ are functions that satisfy Itô conditions \eqref{bc_cond1} and \eqref{bc_cond2}.
 
 One of the main concepts in optimal stopping problems are $optimal stopping times$, which according to \cite{oksendal:book}, we defined as:
 \begin{defi}
	 A function $\tau:\Omega \rightarrow [0,\infty]$ is called a stopping time with respect to the filtration $\mathcal{F}$ is $\{ \omega: \ \tau(\omega)\leq t\} \in \mathcal{F}_t \ \forall t\geq0$.
 \end{defi}
 We will denote $\mathcal{T}$ to be the set of all $\{\mathcal{F}_t\}-$stopping times.
 

 Now the cost/reward function $J$ depends on a state $x\in\mathds{S}$ and a stopping time (instead of a control function) and is such that
 \begin{equation}
 J(x,\tau)=\mathds{E}\left[ \int^\tau_0 \left(e^{-\gamma s} g(X_s) \ ds +e^{-\gamma \tau}h(X_\tau)\right) \mathds{1}_{ \{\tau< \infty \}} \right]  
 \footnote{Recall that we are denoting $\mathds{E}\left[\ . \ | X_0=x\right]$ by $\mathds{E}\left[ \ . \ \right].$},
 \label{bc2_J}
 \end{equation}
 where $g$ denotes a running function and $h$ a terminal function.
 
The value function $V$ depends solely on a state $x\in\mathds{S}$ and it is such that
$V(x)=\inf_{\tau \in \mathcal{S}} J(x, \tau)$, if $J$ is a cost function, and $V(x)=\sup_{\tau \in \mathcal{S}} J(x, \tau)$, if $J$ is a reward function.

Accordingly to our goal, we want to find (or characterize) an optimal stopping time $\tau^\mathcal{T}$ that satisfies $J(x,\tau^*)=V(x) \forall x \in \mathds{R}$. We will see that $\tau^*$ can be defined as a complicated functional of the sample paths of $X$. In order to do so, we split our state space in two regions: a continuation region $\mathcal{C}$ and a stopping region $\mathcal{S}$. Taking their names straightforward, an optimal strategy is the one in which we \textit{continue until its optimal to stop}. Therefore the optimal stopping time is defined as
\begin{equation}
\tau_x^*=\inf\{ t \geq0: \ X^x_t \notin \mathcal{C} \}
=\inf\{ t \geq0: \ X^x_t \in \mathcal{S} \},
\label{stop}
\end{equation}
where we highlight the dependence on the initial state $x$ chosen. Observe that the continuation region $\mathcal{C}$ as a similar job as a Markov control function on optimal control problems, since it represents the strategy to follow.
Considering $J$ to be a cost function, the optimal strategy consists in continuing while the value function $V$ is smaller than the terminal function $h$. Hence the continuation region is given by $\mathcal{C}=\{ x\in \mathds{R}: V(x)<h(x) \}$.


\subsection{Hamilton-Jacobi-Bellman Equation}

In this section we will give an heuristic motivation to the Hamilton-Jacobi-Bellman equation which, in the context of optimal stopping problems, is the equivalent to the dynamic programming equation.

Define $\tilde{\tau}= \inf \{ t \geq \tau: X_t \notin \mathcal{C} \}$ to be a stopping time for a fixed initial state $x \in \mathds{R}$ and an arbitrary $\tau \in \mathcal{T}$.
Observe that the stopping rule of $\tilde{\tau}$ - it is optimal to stop after $\tau$, but might not be optimal before - is similar to the idea of dynamic programming principle, by optimizing with respect to different stopping times $\tau$.

Considering the case where $J$ is a cost function, we obtain by the definition of $V$ that
\begin{align}
V(x)&\leq J(x,\tilde{\tau}) \nonumber  \\
&= \mathds{E}\left[ \int^{\tilde{\tau}}_0 \left(e^{-\gamma s} g(X_s) \ ds +e^{-\gamma \tilde{\tau}}h(X_{\tilde{\tau}}) 
\right) \mathds{1}_{ \{\tilde{\tau}< \infty \}} \right] \nonumber \\
&= \mathds{E} \left[ \int^\tau_0 e^{-\gamma s} g(X_s) \ ds +e^{-\gamma \tau} \left( \int^{\tilde{\tau}}_\tau e^{-\gamma (s-\tau)} g(X_s) \ ds + e^{-\gamma (\tilde{\tau}-\tau)} h(X_{\tilde{\tau}} )  \right) 
\mathds{1}_{ \{\tilde{\tau}< \infty \}} 
\right] \nonumber \\
&=  \mathds{E}\left[ \int^{\tau}_0 e^{-\gamma s} g(X_s) \ ds +e^{-\gamma \tau} J(X_\tau,\tau^*_{X_\tau}) \right] \nonumber \\
&= \mathds{E}\left[ \int^{\tau}_0 e^{-\gamma s} g(X_s) \ ds +e^{-\gamma \tau}V(X_\tau) \right].
 \label{bc2_unf1}
\end{align}

For an arbitrary $\tau\in \mathcal{T}: \ \tau \leq \tilde{\tau}^x$, we have $\tilde{\tau}=\tau_x^*$ and hence the following equality holds
\begin{equation}
V(x)=J(x,\tau_x^*)=J(x,\tilde{\tau})=\mathds{E}\left[ \int^{\tau}_0 e^{-\gamma s} g(X_s) \ ds +e^{-\gamma \tau}V(X_\tau) \right].
	\label{bc2_unf2}
\end{equation}

The dynamic principle for optimal stopping is obtained by combining \eqref{bc2_unf1} with \eqref{bc2_unf2} and it is given by
\begin{equation}
V(x)=\inf_{\tau \in \mathcal{T}} \mathds{E}\left[ \int^{\tau}_0 e^{-\gamma s} g(X_s) \ ds +e^{-\gamma \tau}V(X_\tau) \right].
	\label{bc2_dpp}
\end{equation}

The dynamic programming equation will be now derived following a similar approach as done for stochastic control problems. Consider a fixed $x \in \mathds{R}$ and $\tau \in \mathcal{T}$. Assumming that $e^{-\gamma t} V(x) \in C^2(\mathds(R^2))$, we use Itô's lemma, to integrate it, (as similarly done in \eqref{bc_dpe1}); we take the expectation on both sides (as similarly done in \eqref{bc_ded4}) and we sum the term $\mathds{E}\left[ \int^\tau_0 e^{-\gamma t}g(X_t)dt \right]$ on both sides, obtaining 
\begin{equation}
\mathds{E} \left[e^{-\gamma \tau} + V(X_\tau) \int^\tau_0 e^{-\gamma t}g(X_t)dt \right]=
V(x)+\mathds{E} \left[ \int^t_0  e^{-\gamma s} \left( - \gamma V(X_s))+\mathcal{A}V(X_s)+g(X_s) \right) \ ds \right].
\label{bc2_unf3}
\end{equation}

Since $\tilde{\tau}$ is an optimal stopping time and $\tau< \tau^*_x$, we have that \begin{equation}
V(x)=J(x,\tau^*_x)=J(x,\tilde{\tau}) \  \underset{\eqref{bc2_unf1}}{\Rightarrow} \  V(x)=\mathds{E} \left[e^{-\gamma \tau} + V(X_\tau) \int^\tau_0 e^{-\gamma t}g(X_t)dt \right].
\label{bc2_unf4}
\end{equation} 
Combining deduced results \eqref{bc2_unf3} and \eqref{bc2_unf4}, it follows that
$$\mathds{E} \left[ \int^t_0  e^{-\gamma s} \left( - \gamma V(X_s))+\mathcal{A}V(X_s)+g(X_s) \right) \ ds \right]=0.$$

Assume the continuation region $\mathcal{C}$ to be an open set. Let $x\in \mathcal{C}$ and $\tau_x^*>0$. By dividing last equation for $\tau$ and taking the limit to 0 ($\tau \rightarrow 0$) we obtain
\begin{equation}
- \gamma V(x))+\mathcal{A}V(x)+g(x)=0.
\label{bc2_continuation}
\end{equation}

On the other side, considering $x\notin \mathcal{C}$ (and $\tau_x^*>0$), we have that the value function $V$ is equal to the terminal cost $h$, that is,
\begin{equation}
 V(x)+\mathcal{A}V(x)+g(x)=0.
\label{bc2_stop}
\end{equation}

Considering an arbitrary stopping time $\tau \in \mathcal{T}$, by \eqref{bc2_unf1}, using a similar approach as the one use to deduce \eqref{bc2_continuation} and the definition of the continuation region, we obtain 
\begin{equation}
	\begin{cases}
	- \gamma V(x))+\mathcal{A}V(x)+g(x) \geq 0, \ &x \in \mathcal{C}\\
	h(x)-V(x), \ &x \notin \mathcal{C}
	\end{cases}
	\label{bc2_asd}
\end{equation}

The dynamic programming equation for optimal stopping problems follows by combining the last results, \eqref{bc2_continuation}, \eqref{bc2_stop} and \eqref{bc2_asd}, and corresponds to:
\begin{equation}
\min \{ - \gamma V(x))+\mathcal{A}V(x)+g(x), h(x)-V(x)\}=0, \ x\in \mathds{R}.
\label{bc2_dpe2}
\end{equation}
It is also called \textit{Hamilton-Jacobi-Bellman (HJB) equation}. However instead of an equation, \eqref{bc2_dpe2} is in fact a variational inequality. Also, note that it does not depend explicitly on the continuation, but since, for any state $x$ it recommends us whether to stop or continue, we can deduce the continuation region from which - showing that our previous guess was right.

Observe that all the previous deduction was made considering $J$ to be a cost function. However in the context of this thesis, $J$ will be considered to be a reward function. A similar construction as the one already done could be made, but here we will just state the main results that are going to be used.

If $J$ is a reward function, then we obtain that our optimal stopping problem is such that
\begin{equation}
V(x)=\sup_{\tau \in \mathcal{T}} J(x,\tau)=\sup_{\tau \in \mathcal{T}} \mathds{E}\left[ \int^\tau_0 \left(e^{-\gamma s} g(X_s) \ ds +e^{-\gamma \tau}h(X_\tau)\right) \mathds{1}_{ \{\tau< \infty \}} \right]  
\footnote{Recall (again) that we are denoting $\mathds{E}\left[\ . \ | X_0=x\right]$ by $\mathds{E}\left[ \ . \ \right].$},
\label{stopprob}
\end{equation}
where $g$ corresponds to the running cost function and $h$ to the terminal function.

The associated HJB equation, as deduced in \eqref{bc2_dpe2}, takes the form of
\begin{equation}
\max \{ - \gamma V(x))+\mathcal{A}V(x)+g(x), h(x)-V(x)\}=0, \ x\in \mathds{R},
\label{HJB}
\end{equation}
from which that we deduce that the associated continuation and stopping regions are respectively given by
\begin{align}
\mathcal{C}&=\{ x\in \mathds{R}: h(x)-V(x)<0 \}, \label{contreg}\\
\mathcal{S}&=\{ x\in \mathds{R}: h(x)-V(x)\geq 0 \}=\mathds{R}\setminus \mathcal{C}. \label{stopreg}
\end{align}
The optimal stopping time can be formally defined for any state $x\in \mathds{R}$ as
\begin{equation}
\tau^*_x=\inf \{ \tau \geq 0: \ x\notin \mathcal{C} \}.
\label{stoptime}
\end{equation}

\subsubsection{Verification}
As well as done in optimal control problems, a solution $\phi$ is found using the verification technique now applied on the context of optimal stopping problems.

We won't go further in detail since the methodology is pretty similar to what was described before. Instead we will give the main idea of it. For more information one should check \cite{ross} and \cite{oksendal:book}. 

As the verification technique was previously explained, it doesn't give a solution by solving explicitly the HJB variational inequality \eqref{HJB}, but instead, we assume a solution $\phi$ and we show that it verifies the HJB variational inequality.

The function $\phi$ might be defined by parts, in order to verify both sides of the variational inequality \eqref{HJB}. However the main difficulty appears at on the boundary of $\mathcal{C}$. The strong formulation of the verification technique - as stated on Section \ref{section:control} - requires $\phi$ to be $C^2$. This assumption is hardly verified since there is an abrupt change on the behaviour from the $\mathcal{C}$ to $\mathcal{S}$. Fortunately, there are generalizations of Itô's rule that only require $\phi$ to be $C^1$ (and other weaker assumptions). This condition over $\phi$ will originate two other conditions referred as \textit{value matching} and \textit{smooth pasting} conditions, as stated in \cite{dixit:book}, essential to the formulation of $\phi$.

Therefore, having $phi$ constructed by parts such that it verifies the HJB variational inequality in \eqref{HJB} and is $C^1$ everywhere, we have found the solution to the optimal stopping problem \eqref{stopprob}.





\subsection{Optimal Stopping Problems Applied to Investment Decisions under Uncertainty}

In this section we will deduce the solution to all standard optimal stopping problems that we will face during this thesis, taking into account the respective financial context.

We will start by presenting some financial concepts related to investment decisions under uncertainty, based on \cite{dixit:book}.
Then we will return to the mathematical formulation and derive the solution, presenting a detailed explanation.

\subsubsection{Investment Decisions under Uncertainty}

Quite many recent works suggest that a Real Option analysis is much more advantageous than the Net Present Value (NPV) analysis, since the last one assumes that the company has a now or never approach regarding the investment decision, since it can only exercise at the moment of analysis. On the other side a Real Option Analysis takes into account the future potential, with the respective uncertainty.

The first contributions on the Real Options approach were due to McDonald \& Siegel(1986) \cite{siegel}, in whose work they model an investment problem where the investor must decide when it is the best time to exercise, taking into account that the value of the investment project is stochastically random, evolving accordingly to a Geometric Brownian Motion, and Dixit (1989) \cite{dixit_alone}, in whose work he model the best time to make entry and exit decisions, while considering that there might be costs associated to each them, and taking into account that the market price evolves accordingly to a Brownian Motion.

Years later, in 1994, Dixit \& Pindyck created \cite{dixit:book}, which exploits an analogy between real options and financial investment decisions, focusing on many different decision problems (entry, investment and exit, among them) dependent on different stochastically behavioured (diffusion processes and jump diffusion processes, among them) measures, such as demand or market price. This book is considered by some of the experts as the (financial) \textit{Bible} of Real Options approach. 

More recently, Hagspiel \textit{et al.} (2016) \cite{hagspiel:cap} explores the case in which a firm must decide to invest in a new product (or exit the market) and when it must definitely exit its production, while facing a declining profit for the established product. The solution leads to three different thresholds, based on the (stochastic) demand level, for the respective possible decisions.

In this work we will also consider the demand to be stochastically, evolving accordingly to a Geometric Brownian Motion. On the third optimal stopping problem, on Section \ref{chapter:3}, we will also derive three different thresholds, although these are related with the possibility of either invest in a new product with simultaneous production of the new and the established product and then stop the production of the established product or invest on the new product and immediately replace the established product. A similar situation was already presented by Pimentel \cite{rita}, however in a different context. On her work, she considers two sources of uncertain - which strongly influence each of the thresholds derived -, while in our work we only consider one, the demand level already referred. 


