%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Introduction.tex                                    %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Andre C. Marta                                           %
%     Last modified :  2 Jul 2015                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}
\label{chapter:introduction}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
\label{section:motivation}

Nowadays, society is highly attached to technology, from mobile phones to fancy gadgets - and sometimes scratching the absurd
\footnote{Parents forget to feed child while playing videogames:\\
\url{https://www.theguardian.com/world/2010/mar/05/korean-girl-starved-online-game} }.
 Nevertheless, all this huge demand brought changes on how IT companies should manage their investments. These should, not only pay attention to the product demand on the market, but also to technology evolution. Therefore investors began to require more complex models to support their decisions. Models must not only consider the current value of the firm, such as the ones based on the Net Present Value (NPV), but also must take into account the potential associated to future events.

As an example ASML, a dutch company considered to be the largest european semiconductor equipment maker which currently is part of Euro Stoxx 50\footnote{Stock index representing the 50 most liquid stocks in the Eurozone:\\  \url{http://en.boerse-frankfurt.de/index/constituents/Euro_Stoxx_50\#Constituents} }.
ASML builds electronic chips producing machines that satisfy the needs of hardware building companies - for instance, Intel and Samsung\footnote{For further details check: \url{https://www.asml.com/company/our-history/en/s277?rid=51985}}.
Therefore, in order to survive on the competitive market, ASML needs either to develop their own technology, by investing on R\&D, or to import a desired one, with an associated cost.

This work is focused on the case where the firm chooses to develop their own technology and, hence, define \textit{a priori} the desired innovation level and investment to be made. Since it is impossible to access that a certain level of technology will be reached in a precise amount of time\footnote{Flying cars predicted for 2015 in \textit{Back to Future} as an example (!):\\  \url{http://content.time.com/time/specials/packages/article/0,28804,2024839_2024845_2024855,00.html}} \footnote{Moore's Law - which defends that processor performance doubles (approximately) every two years - is dying:\\
\url{https://www.forbes.com/sites/forbestechcouncil/2018/03/09/moores-law-is-dying-so-where-are-its-heirs/\#53cb24a17a7b}}, the innovation process is considered to evolve randomly with time, with a rate that might be influenced by the amount of money invested. More money implies more resources and, hence, an higher evolution rate.

After %having reached 
reaching the desired innovation level, the firm must evaluate under which market conditions it should invest and in which type of products. The final goal is to maximize the expected long run profit.


%iPhone X release in 2017:\\ \url{https://www.bloomberg.com/news/articles/2017-11-02/long-lines-are-back-at-apple-stores-for-10th-anniversary-iphone
%}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem Context}
\label{section:context}

In the current work we consider a firm that develops its own technology and decides to invest in a new product only after an innovation breakthrough, associated to innovation level $\theta$.


The innovation process associated to the technology level is denoted by $\{ \theta(t), \ t \geq 0 \}$ and considered to be an homogeneous Compound Poisson process, evolving accordingly to
$$\theta_t= \theta_0+ u N_t$$  
where $\theta_0$ corresponds to the initial innovation level, $u > 0$ to the jump size in the technology and $\{N_t, \ t \geq 0\}$  to the jump process, which follows a Poisson process with rate $\lambda(R)=R^\gamma, \ \gamma \in (0,1)$, where $R$ stands for the R\&D investment, such as scientist salaries, equipment costs and expenses. The mentioned rate function $\lambda$ is such that it verifies:
\begin{itemize}
	\item $\lambda(0) = 0$: no R\&D investment means zero probability of innovating;
	\item $\lambda^\prime (R)>0$: the larger the investment is, the higher is the probability of observing a jump. Hence, the waiting time for the next technology jump is expect to be shorter;
	\item $ \lambda ^{ \prime \prime} (R)<0$: exists an optimal R\&D investment that leads to the maximization of the rate function, that is, $\exists R^*: \lambda(R^*)\geq \lambda(R), \  \forall R$.
\end{itemize}

%Under this situation the firms decides how much it will invest in R\&D, knowing that a bigger investment implies a smaller expected time until it is able to invest on the new product.

When the breakthrough happens, the innovation process stops and the firm is able to decide when it should invest on the new product with the innovation level $\theta$. In order to have the new product available on the market, the firm needs to incur an investment (sunk) cost. This one is related with employees' formation and new equipment, for example, that depends on the production capacity chosen $K_1$. In this work, it is assumed that the firm produces up to its full capacity. Hence sunk costs will be dependent on the quantity of production chosen.

Since the new product offers a new technology level, not recognized by the market, its profit is strongly dependent on market's demand. The better the reception of the product, the bigger the demand and the corresponding profit.
%Here the demand is considered to evolve accordingly to a Geometric Brownian Motion (GBM) which initial state corresponds to the demand at the breakthrough time.

%The investment decision will be influenced by the long run profit possible to be obtained.

The goal of this work is to define,
%using realoptions analysis,
 the optimal investment time on a new product. The estimation of this time takes into account Real Options analysis, how market conditions influences it and 
 %hat 
 the maximization of the expected long run profit.
 % is maximized and to study how market conditions influence it.

In the following sections we will consider the timeline represented in the schematic diagram below 
\colorbox{red}{pôr em Figura e fazer  novas figuras > TIF}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%\begin{tikzpicture}[y=1cm, x=1cm, thick, font=\footnotesize]    
%\usetikzlibrary{arrows,decorations.pathreplacing}
%
%\tikzset{
%	brace_top/.style={
%		decoration={brace},
%		decorate
%	},
%	brace_bottom/.style={
%		decoration={brace, mirror},
%		decorate
%	}
%}
%
%% time line week
%\draw[line width=1.2pt, ->, >=latex'](0,0) -- coordinate (x axis) (14,0) node[right] {\textit{Time}}; 
%\foreach \x in {0,3,8}
%\draw (\x cm,3pt) -- (\x cm,-3pt);
%%\foreach \x in {current time,  , $\tau$, 3} \draw (\x,0.1) -- (\x,-0.1) node[below] {\x};
%\draw (0,0) node[below=3pt] (a) {$\text{R\&D investment}$} node[above=3pt] {};
%\draw (3,0) node[below=9pt] (a) {$\text{Innovation breakthrough $\theta$\\
%	$t_1=0$}$} node[above=3pt] {};
%\draw (8,0) node[below=3pt] (a) {$\tau:$ 
%\text{ \textit{new} product's investment\\ $t_2=0$: initial instant regarding $\tau_2$}} node[above=3pt] {};
%\draw (12,0) node[below=9pt] (a) {$\tau_2:$ \text{\textit{old} product's abandonment}} node[above=3pt] {};
%
%%brace
%%\node (start_p) at (0,0.6) {};
%%\node (end_p) at (7.99,0.6) {};
%%\draw [brace_top] (start_p.north) -- node [above, pos=0.5] {$\text{\textit{Old} product}$} (end_p.north);
%
%%brace
%%\node (start_pn) at (8.01,0.6) {};
%%\node (end_pn) at (11.99,0.6) {};
%%\draw [brace_top] (start_pn.north) -- node [above, pos=0.5] {$\text{\textit{Old \& New} product}$} (end_pn.north);
%
%%brace
%\node (start_pn) at (8.01,0.6) {};
%\node (end_pn) at (13.8,0.6) {};
%\draw [brace_top] (start_pn.north) -- node [above, pos=0.5] {$\text{\textit{New} product}$} (end_pn.north);
%
%
%% top brace
%\node (start_t) at (0,0.1) {};
%\node (end_t) at (2.99,0.1) {};
%\draw [brace_top] (start_t.north) -- node [above, pos=0.5] {$\text{Innovation development}$} (end_t.north);
%
%% top brace
%\node (start_week) at (3.01,0.1) {};
%\node (end_week) at (13.8,0.1) {};
%\draw [brace_top] (start_week.north) -- node [above, pos=0.5] {$\text{Firm is able to make its investment decisions regarding $\theta$}$} (end_week.north);
%
%\end{tikzpicture}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{4mm}

\begin{tikzpicture}[y=1cm, x=1cm, thick, font=\footnotesize]    
\usetikzlibrary{arrows,decorations.pathreplacing}

\tikzset{
	brace_top/.style={
		decoration={brace},
		decorate
	},
	brace_bottom/.style={
		decoration={brace, mirror},
		decorate
	}
}

% time line week
\draw[line width=1.2pt, ->, >=latex'](0,0) -- coordinate (x axis) (14,0) node[right] {\textit{Time}}; 
\foreach \x in {0,3}
\draw (\x cm,3pt) -- (\x cm,-3pt);
%\foreach \x in {current time,  , $\tau$, 3} \draw (\x,0.1) -- (\x,-0.1) node[below] {\x};
\draw (0,0) node[below=3pt] (a) {$\text{R\&D investment}$} node[above=3pt] {};
\draw (3,0) node[below=9pt] (a) {$\text{Innovation breakthrough $\theta$}$} node[above=3pt] {};

%%brace
%\node (start_p) at (0,0.6) {};
%\node (end_p) at (7.99,0.6) {};
%\draw [brace_top] (start_p.north) -- node [above, pos=0.5] {$\text{\textit{Old} product}$} (end_p.north);
%
%%brace
%\node (start_pn) at (8.01,0.6) {};
%\node (end_pn) at (11.99,0.6) {};
%\draw [brace_top] (start_pn.north) -- node [above, pos=0.5] {$\text{\textit{Old \& New} product}$} (end_pn.north);
%
%%brace
%\node (start_pn) at (12.01,0.6) {};
%\node (end_pn) at (13.8,0.6) {};
%\draw [brace_top] (start_pn.north) -- node [above, pos=0.5] {$\text{\textit{New} product}$} (end_pn.north);


% top brace
\node (start_t) at (0,0.1) {};
\node (end_t) at (2.99,0.1) {};
\draw [brace_top] (start_t.north) -- node [above, pos=0.5] {$\text{Innovation development}$} (end_t.north);

% top brace
\node (start_week) at (3.01,0.1) {};
\node (end_week) at (13.8,0.1) {};
\draw [brace_top] (start_week.north) -- node [above, pos=0.5] {$\text{Firm is able to make its investment decisions regarding $\theta$}$} (end_week.north);

\end{tikzpicture}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Henceforth, when determining the optimal investment time on the new product, we consider as the starting time the instant at which occurs the innovation breakthrough. However when evaluating the value function regarding the complete investment (associated to both R\&D and new product), the starting time corresponds to the time at which the firm decides to invest in R\&D.





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{State-of-the-art}

%Quite many recent works
%suggest that Real Option\footnote{Real Option's concept will be defined on Section \ref{bc_ro}.} analysis is much more advantageous than the Net Present Value (NPV) analysis when it comes to support investment decisions, since the last one assumes that the company has a now or never approach regarding the investment decision, ignoring the freedom in timing.
%and hence it ignores the option of having a better situation in the future. 
%On the other side a Real Option Analysis takes into account the future potential, as well the respective uncertainty.
%Among many studies already made, Farzin \textit{et al.} (1998) \cite{farzin:cap}, do a comparative study between both approaches in the technology adoption context.

One of the first contributions on Real Options analysis regarding investment decisions was due to McDonald \& Siegel (1986)  \cite{siegel}. They model an investment problem where the investor must decide when it is the best time to exercise the option, taking into account that the value of the investment project is stochastically random and evolves accordingly to a Geometric Brownian Motion. Other contributions were due to Dixit (1989) \cite{dixit_alone}, who models the best time to make entry and exit decisions, while considering that the market price evolves accordingly to a Brownian Motion and that each decision has an associated cost.

The study by Dixit \& Pindyck (1994) \cite{dixit:book}, that is nowadays considered the financial by some Bible of Real Options approach, exploits an analogy between real options and financial investment decisions, focusing on many different decision problems (entry, investment and exit, among them) dependent on different stochastically behavioured (diffusion processes and jump diffusion processes, among them) measures, such as demand or market price.

%Some other topics regarding Real Option increased their relevance. More particularly, the optimal production capacity to be chosen and the impact of technology adoption - being this last one a field of increasing interest nowadays.

As our understanding about the market of new technological products evolved, the optimal production capacity to be chosen and the impact of technology adoption associated to investment decisions have started to get more relevant.

%Regarding optimal production capacity, 
In the aspect of optimal capacity production Huisman \& Kort (2013) \cite{huis:cap} proposed a methodology to estimate the best time and capacity to invest in a new product. Both monopoly and a duopoly situations were assumed on their study.

Regarding technology adoption,
Farzin \textit{et al.} (1998) \cite{farzin:cap}, presented a comparative study between NPV and Real Option approaches in the context that a firm wants to deduce when is the best time and level to invest in a technology.
Also, Hagspiel \textit{et al.} (2016) \cite{hagspiel:cap} studied the best time to invest in a new product or exit the market, by considering a firm whose established product has a declining profit and a demand process to evolve accordingly to a GBM. %The solution leads to three different demand thresholds for the respective possible decisions.


Recently, Pimentel (2018) \cite{rita} explored both optimal capacity level and technology adoption. By considering two sources of uncertainty, related to the demand and the innovation processes,
%- the demand to evolve as a jump diffusion process and the innovation process as a compound Poisson process - 
and a firm which is producing an established product, she deduces the optimal times to invest in a new product and to stop the production of the established product. 
%Regarding the described situation, two models were developed: the benchmark model and the capacity optimization model.



%In this work we will also consider the demand to be stochastically, evolving accordingly to a Geometric Brownian Motion. On the third optimal stopping problem, on Section \ref{chapter:3}, we will also derive three different thresholds, although these are related with the possibility of either invest in a new product with simultaneous production of the new and the established product and then stop the production of the established product or invest on the new product and immediately replace the established product. A similar situation was already presented by Pimentel \cite{rita}, however in a different context. On her work, she considers two sources of uncertain - which strongly influence each of the thresholds derived -, while in our work we only consider one, the demand level already referred. 




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Thesis Outline}
\label{section:outline}


This thesis is organized in eight chapters.

The first one consists in the present introduction, which includes the state-of-the-art the context considered along this thesis.

%Considering the context described on Section \ref{section:context}, we will treat three different situations related with the irreversible decision of investing in a new product when the innovation breakthrough and future demand are stochastic.

%\textcolor{red}{1 SOURCE OF UNCERATINTY (DEMAND) OU 2 (INNOVATION PROCESS)?}

The major theoretical concepts and the state-of-the-art are introduced on Chapter \ref{chapter:bc}, following several approaches presented in the literature (\cite{dixit:book}, \cite{ross} and \cite{oksendal:book}).
We start by exploring the most relevant results of Optimal Stopping Problems (OSP). Then we show how they are related to investment decisions under uncertainty through Real Option analysis.

On the next three chapters we analyse different possible investment decisions, namely:
\begin{itemize}
	\item Chapter \ref{chapter:1}: a firm that has no products in the market and wants to find the optimal time to introduce a new product;
	\item Chapter \ref{chapter:2}: a firm with an established product in the market wants to find the optimal time to invest in a new product while replacing immediately the old one;
	\item Chapter \ref{chapter:3}: a firm with an established product wants to find the optimal time to invest in a new product.
	Two scenarios are considered:
	 \begin{enumerate}
	 	\item The old product is immediately replaced;
	 	\item A simultaneous production period is allowed and followed by the removal of the oldest product.
	 \end{enumerate}
	% while either replacing immediately the established product by the new one or allowing a simultaneous production period, followed by the removal of the oldest product.
\end{itemize}

conditions

Two models are developed for situations described on Chapters \ref{chapter:1} and \ref{chapter:2}. The first one corresponds to the benchmark model, which considers the original cash-flow for a chosen production capacity. The second one to the capacity optimization model, which considers the maximized long-run cash-flow with respect to production capacity. 
Regarding the situation described on Chapter \ref{chapter:3}, due to its complexity, only the benchmark model is analysed.
On the last section of each of the three chapters we study the behaviour of respective thresholds and optimal capacity level with different parameters.


The behaviour of optimal investment times, associated to the situations described on chapters 3 to 5, regarding initial demand values and market's uncertainty is studied on Chapter \ref{chapter:stoptime}.



On Chapter \ref{chapter:max} we derive the optimal R\&D investment, by maximizing the expected value function with respect to the innovation process. 
First, we derive the optimal R\&D investment considering that the innovation process only takes one jump to achieve the breakthrough level.
Secondly, we generalize the previous situation by considering that the innovation process takes $n \in \mathds{N}$ jumps to achieve the breakthrough level.
In the end of this chapter, the behaviour of the decision threshold with the different parameters is studied.



Finally, on Chapter \ref{chapter:conc} we summarize the relevant findings and how this work can be extended.


